---
title: "Reproducible document"
author: "Alexander"
format: html
---

## Basics of R

```{r}
1 + 1
```
## Setup Code Chunk
```{r setup}
#| message: false
#| warning: false
library(tidyverse)
post_meal_data <- read_csv2(
  here::here("data/post-meal-insulin.csv")
)

```
## About me
- Alexander Hinz
- SDU
I am a new **Ph.d Student** at FGM under the supervision by *Susanne* Mandrup. My work is focused on in vivo adipogenesis in humans adipose tissue.

## Simple code
```{r}
3 * 3
```

## Testing for Git
Its cold in Denmark at this time of the year

## Showing the data
```{r}
post_meal_data
```

```{r}
glimpse(post_meal_data)
```
## Visualization of Data
```{r}
ggplot(post_meal_data, aes(x = BMI)) + geom_histogram()
```
```{r}
#| fig-cap: "Distribution of BMI"
#| label: fig-bmi-histo
ggplot(post_meal_data, aes(x = BMI)) + geom_histogram()
```

Making a figure link that opens a windows figure when hovering over it. 
@fig-bmi-histo

## Exercise discerete plots

```{r}
#| fig-cap: "Discrete plotting of glykemi"
#| label: fig-glykemi

ggplot(post_meal_data, aes(glykemi)) + geom_bar()

```

@fig-glykemi

