---
title: "Reproducible document"
author: "Alexander"
format: html
---

## Basics of R

```{r}
1 + 1
```
## Setup Code Chunk
```{r setup}
#| message: false
#| warning: false
library(tidyverse)
post_meal_data <- read_csv2(
  here::here("data/post-meal-insulin.csv")
)

small_post_meal_data <- select(post_meal_data, id, Age, BMI, Group, auc_pg, auc_ins, glykemi)
```
## About me
- Alexander Hinz
- SDU
I am a new **Ph.d Student** at FGM under the supervision by *Susanne* Mandrup. My work is focused on in vivo adipogenesis in humans adipose tissue.

## Simple code
```{r}
3 * 3
```

## Testing for Git
Its cold in Denmark at this time of the year

## Showing the data
```{r}
post_meal_data
```

```{r}
glimpse(post_meal_data)
```
## Visualization of Data
```{r}
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```
```{r}
#| fig-cap: "Distribution of BMI"
#| label: fig-bmi-histo
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

Making a figure link that opens a windows figure when hovering over it. 
@fig-bmi-histo

## Exercise discerete plots

### Plotting one discrete values
```{r}
#| fig-cap: "Discrete plotting of glykemi"
#| label: fig-glykemi

ggplot(post_meal_data, aes(glykemi)) +
  geom_bar()
```

@fig-glykemi

### Plotting two discrete values
```{r}
post_meal_data |>
  ggplot(aes(x = Group, fill = glykemi)) +
  geom_bar()
```

```{r}
post_meal_data |>
  ggplot(aes(x = Group, fill = as.character(glykemi))) +
  geom_bar()
```

Position_dodge() places information next to each other
```{r}
post_meal_data |>
  ggplot(aes(x = Group, fill = as.character(glykemi))) +
  geom_bar(position = position_dodge())
```
## Putting two plots side by side

```{r BMI histogram}
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

```{r glykemi barplot}
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```

```{r side-by side-figs}
#| label: fig-bmi-glykemi
#| fig-cap: "BMI and glykemi, side by side"
#| fig-subcap:
#| - "Distribution of BMI"
#| - "Number of those with glykemi"
#| layout-ncol: 2

ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()

ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```

## Plotting two continous variables
```{r scatterplot BMI and plasma glucose}
#| label: fig-bmi-auc-pc
#| fig-cap: "Scatterplot of correlation between BMI and Glucose"

ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point() +
  geom_smooth()
```
## Pushing changes to remote
## Pulling changes to local PC

## Manipulating Data
select(dataset, variable of interest). "-" sign indicates removing column.
starts_with("") extracts data that has a specific character or set. Also works with ends_with("")
contains("") extracts information in the dataset where the specific set of characters occurs.
```{r Selecting Data}
select(post_meal_data, Age)

select(post_meal_data, Age, BMI, Weight)

select(post_meal_data, -Group)

select(post_meal_data, starts_with("PG"))

select(post_meal_data, ends_with("DXA"))

select(post_meal_data, contains("0"))
```
